---
title: "TSA: Forecasting Competition "
author: "Yuxiang Ren and Karen Thornton"
output: pdf_document
always_allow_html: true
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---
#packges    
```{r packages}
library(readxl)
#library(xlsx) #this package doesn't work on Macs very well
library(tidyr)
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)

library(tidyverse)
library(smooth)
library(zoo)
library(kableExtra)
library(dplyr)

```

#Data
##temperature
```{r}
#temperature
raw_temperature <- read_excel("Competition/Data/temperature.xlsx")
colnames(raw_temperature) 

##transfer hourly to daily
temperature_day <- raw_temperature %>% 
  group_by(date) %>% 
  summarize(across(starts_with("t_ws"), mean, na.rm = TRUE))
##test result use first day of t_ws1
mean(raw_temperature$t_ws1[c(1:24)])
##save
write.csv(temperature_day, row.names = FALSE, 
          file = "./Competition/Data/temperature_day.csv")
```
## load
```{r}
#load
raw_load <- read_excel("Competition/Data/load.xlsx")
colnames(raw_load) 
## i guess h1 ~h24 mean hours,
## transfer hourly to daily
load_day_long <- raw_load %>%
  pivot_longer(cols = starts_with("h"),
               names_to = "hour",
               values_to = "energy")
load_day <- load_day_long %>% 
  group_by(date) %>% 
  summarize(energy = mean(energy))# use mean

#changed to mean 

##test
sum(load_day_long$energy[1:24])

##save
write.csv(load_day, row.names = FALSE, 
          file = "./Competition/Data/load_day.csv")
```

##humidity
```{r}
#humidity
raw_humidity <- read_excel("Competition/Data/relative_humidity.xlsx")
colnames(raw_humidity) 
##transfer hourly to daily
humidity_day <- raw_humidity %>% 
  group_by(date) %>% 
  summarize(across(starts_with("rh_ws"), mean, na.rm = TRUE))
##test result use first day of t_ws1
mean(raw_humidity$rh_ws1[c(1:24)])
##save
write.csv(humidity_day, row.names = FALSE, 
          file = "./Competition/Data/humidity_day.csv")
```

##final data
**to save space i save the result data, you can only run after this chunk to load data. or we can add "#" before these r chunk**
```{r}
load_day <- read.csv("./Competition/Data/load_day.csv")
temperature_day <- read.csv("./Competition/Data/temperature_day.csv")
humidity_day <- read.csv("./Competition/Data/humidity_day.csv")

```

#create a time series
```{r}
ts_load_daily <- msts(load_day$energy, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,1,1))

load_day_fill <- na.approx(load_day$energy)
ts_load_daily_fill <- msts(load_day_fill, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,1,1))

data_seq_YR <- seq(from = as.Date("2011-01-01"), to = as.Date("2011-02-28"), by = "day")

#ts_temperature
##mean temperature
temp <- temperature_day %>%
  group_by(date) %>%
  summarise(T = mean(c_across(starts_with("t_ws")), na.rm = TRUE))

ts_temp <- msts(temp,seasonal.periods =c(7,365.25),
                           start=c(2005,1,1))
acf(load_day_fill)
pacf(load_day_fill)
```

#Test_Karen
##create an arima
```{r}
#arima with only one factor
Simple_ARIMA <- auto.arima(ts_load_daily, seasonal=FALSE)
print(Simple_ARIMA)

#simple arima forecast
Simple_ARIMA_for <- forecast(object = Simple_ARIMA, h=365)

#plot
autoplot(ts_load_daily) +
    autolayer(Simple_ARIMA_for,series="ARIMA",PI=FALSE) +
    ylab("Load") +
    xlab("")

#save
write.csv(Simple_ARIMA_for, row.names = FALSE, 
          file = "./Competition/Data/Simple.ARIMA.Forecast.mean.csv")
```

#Test_Karen_2
##Neural Network
```{r}
NN_fit <- nnetar(ts_load_daily,p=1,P=0,xreg=fourier(ts_load_daily, K=c(2,12)))

NN_for <- forecast(NN_fit, h=59,xreg=fourier(ts_load_daily, 
                                          K=c(2,12),h=59))

#plot
autoplot(ts_load_daily) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  ylab("Load") 

#save
write.csv(NN_for, row.names = FALSE, 
          file = "./Competition/Data/Neural.Network.Forecast.csv")
```

#Test_Karen_3
##Neural Network 2
```{r}
NN_fit_2 <- nnetar(ts_load_daily,p=1,P=0,xreg=fourier(ts_load_daily, K=c(2,7)))

NN_for_2 <- forecast(NN_fit_2, h=59,xreg=fourier(ts_load_daily, 
                                          K=c(2,7),h=59))

#plot
autoplot(ts_load_daily) +
  autolayer(NN_for_2, series="Neural Network 2",PI=FALSE)+
  ylab("Load") 

#save
write.csv(NN_for_2, row.names = FALSE, 
          file = "./Competition/Data/Neural.Network.Forecast.2.csv")
```

#Test_Yuxiang
##STL + ETS
```{r STL+ETS}
ETS_fit_YR <-  stlf(ts_load_daily,h=59)
autoplot(ETS_fit_YR)
STL_ETS_YR <- ETS_fit_YR$mean

#save

data_seq_YR <- seq(from = as.Date("2011-01-01"), to = as.Date("2011-02-28"), by = "day")
STL_ETS_YR_df <- data.frame(date = data_seq_YR, load = STL_ETS_YR)

write.csv(STL_ETS_YR_df, row.names = FALSE, 
          file = "./Competition/Data/STL.ETS.Forecast.csv")

```

##ARIMA + FOURIER terms
loop take long time, the best K values is (2,2)
```{r ARIMA.FOURIER}

min_AICc <- Inf
best_K <- c(0, 0)

for (k1 in 1:2) {
  for (k2 in 1:20) { # Change the range of k2 to a smaller value for testing
    ARIMA_Four_fit_temp <- auto.arima(ts_load_daily,
                                      seasonal=FALSE,
                                      lambda=0,
                                      xreg=fourier(ts_load_daily,
                                                   K=c(k1, k2))
                                      )
    current_AICc <- ARIMA_Four_fit_temp$aicc
    
    cat("k1: ", k1, " k2: ", k2, " AICc: ", current_AICc, "\n")
    
    if (current_AICc < min_AICc) {
      min_AICc <- current_AICc
      best_K <- c(k1, k2)
    }
  }
}

print(paste0("Best K values: (", best_K[1], ", ", best_K[2], ")"))
print(paste0("Minimum AICc: ", min_AICc))

ARIMA_Four_fit_YR <- auto.arima(ts_load_daily, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_load_daily, 
                                          K=best_K)
                             )

ARIMA_Four_for_YR <- forecast(ARIMA_Four_fit_YR,
                           xreg=fourier(ts_load_daily,
                                        K=best_K,
                                        h=59),
                           h=59
                           ) 

#save
ARIMA_Four_YR <- ARIMA_Four_for_YR$mean
data_seq_YR <- seq(from = as.Date("2011-01-01"), to = as.Date("2011-02-28"), by = "day")
ARIMA_Four_YR_df <- data.frame(date = data_seq_YR, load = ARIMA_Four_YR)

write.csv(ARIMA_Four_YR_df, row.names = FALSE, 
          file = "./Competition/Data/ARIMA.FOUR.Forecast.csv")

#re-test k = 2, 12
ARIMA_Four_fit2_YR <- auto.arima(ts_load_daily, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_load_daily, 
                                          K=c(2,12))
                             )

ARIMA_Four_for2_YR <- forecast(ARIMA_Four_fit2_YR,
                           xreg=fourier(ts_load_daily,
                                        K=c(2,12),
                                        h=59),
                           h=59
                           ) 

#save
ARIMA_Four2_YR <- ARIMA_Four_for2_YR$mean
data_seq_YR <- seq(from = as.Date("2011-01-01"), to = as.Date("2011-02-28"), by = "day")
ARIMA_Four2_YR_df <- data.frame(date = data_seq_YR, load = ARIMA_Four2_YR)

write.csv(ARIMA_Four2_YR_df, row.names = FALSE, 
          file = "./Competition/Data/ARIMA.FOUR2.Forecast.csv")

```

##TBATS
```{r}
TBATS_fit1_YR <- tbats(ts_load_daily)
TBATS_for1 <- forecast(TBATS_fit1_YR, h=59)
autoplot(TBATS_for1) +
  ylab("Active Power") # too small, not report


#replace NA, might wrong
ts_full_YR <- approx(ts_load_daily, method = "linear")
TBATS_fit2_YR <- tbats(ts_full_YR$y)
TBATS_for2 <- forecast(TBATS_fit_YR, h=59)
autoplot(TBATS_for2) +
  ylab("Active Power") 


```

##Neural_Network
same code as your NN2, but set seed 
```{r Neural_Network}
set.seed(6787)
NN_fit_2_YR <- nnetar(ts_load_daily,p=1,P=0,xreg=fourier(ts_load_daily, K=c(2,7)))

NN_for_2_YR <- forecast(NN_fit_2_YR, h=59,xreg=fourier(ts_load_daily, 
                                          K=c(2,7),h=59))

#plot
autoplot(ts_load_daily) +
  autolayer(NN_for_2_YR, series="Neural Network 2",PI=FALSE)+
  ylab("Load") 

#save
Neural.Network.6787.YR_df <- data.frame(date = data_seq_YR, load = NN_for_2_YR$mean)

write.csv(Neural.Network.6787.YR_df, row.names = FALSE, 
          file = "./Competition/Data/Neural.Network.Forecast.seed6787.2.csv")


```

##Neural_Network, k= 2,2
maybe need setseed
```{r Neural_Network(2,2)1}
set.seed(6787)
NN_fit_3_YR <- nnetar(ts_load_daily,p=1,P=0,xreg=fourier(ts_load_daily, K=c(2,2)))

NN_for_3_YR <- forecast(NN_fit_3_YR, h=59,xreg=fourier(ts_load_daily, 
                                          K=c(2,2),h=59))

#plot
autoplot(ts_load_daily) +
  autolayer(NN_for_3_YR, series="Neural Network 3",PI=FALSE)+
  ylab("Load") 

#save
Neural.Network.22.6787.YR_df <- data.frame(date = data_seq_YR, load = NN_for_3_YR$mean)

write.csv(Neural.Network.22.6787.YR_df, row.names = FALSE, 
          file = "./Competition/Data/Neural.Network.Forecast.22.seed6787.csv")


```
##Neural_Network, k= 2,2
seed = 1234
```{r Neural_Network(2,2)2}
set.seed(1234)
NN_fit_4_YR <- nnetar(ts_load_daily,p=1,P=0,xreg=fourier(ts_load_daily, K=c(2,2)))

NN_for_4_YR <- forecast(NN_fit_4_YR, h=59,xreg=fourier(ts_load_daily, 
                                          K=c(2,2),h=59))

#plot
autoplot(ts_load_daily) +
  autolayer(NN_for_4_YR, series="Neural Network 3",PI=FALSE)+
  ylab("Load") 

#save
Neural.Network.22.1234.YR_df <- data.frame(date = data_seq_YR, load = NN_for_4_YR$mean)

write.csv(Neural.Network.22.1234.YR_df, row.names = FALSE, 
          file = "./Competition/Data/Neural.Network.Forecast.22.seed1234.csv")

```
##Neural_Network, find the best pPK.
this code can not work yet
```{r}
#replace NA
load_day_fill <- na.approx(load_day$energy)
ts_load_daily_fill <- msts(load_day_fill, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,1,1))
ts_data_train <- window(ts_load_daily_fill, start=c(2005, 1), end=c(2009, 365))

ts_data_test <- window(ts_load_daily_fill, start=c(2010, 1), end=c(2010, 365))

min_MSE <- Inf
best_p <- 0
best_P <- 0
# Loop through different combinations of p, P, and K
for (p in 0:3) {
  for (P in 0:3) {
        # Fit the model with the current combination of parameters
       NN_fit <- nnetar(ts_data_train, p=p, P=P, xreg=fourier(ts_data_train, K=c(2,12)))
       NN_for_temp <- forecast(NN_fit, h = 365, xreg=fourier(ts_data_test, K = c(2,12), h = 365))

       current_MSE <- mean((ts_data_test - NN_for_temp$mean)^2)

        if (current_MSE < min_MSE) {
          min_MSE <- current_MSE
          best_p <- p
          best_P <- P
        }
      }
    }

cat("Best p:", best_p, "\n")
cat("Best P:", best_P, "\n")
cat("Minimum MSE:", min_MSE, "\n")


```

##Neural_Network, for non-seasonal
i forget there should no k
```{r NN, non-seasonal,k(2,12)}
## remove seansaol
stl_decomposition <- stl(ts_load_daily_fill, s.window = "periodic", robust = TRUE)
seasonal_component <- stl_decomposition$time.series[, "seasonal"]
deseasonalized_data <- ts_load_daily_fill - seasonal_component

set.seed(6787)
NN_fit_deseasonalized_212 <- nnetar(deseasonalized_data, p = 1, P = 0, K = c(2, 12))
NN_forecast_deseasonalized_212 <- forecast(NN_fit_deseasonalized_212, h = 59)

forecast_seasonal_component <- window(seasonal_component, start = c(2010, 1), end = c(2010, 59))


df <- data.frame(date = data_seq_YR, load1 = NN_forecast_deseasonalized_212$mean)
df2 <- data.frame(load2 = forecast_seasonal_component)
df3 <- cbind(df,df2) 
df3$load <- df3$load1 + df3$load2

df <- data.frame(date = as.Date(data_seq_YR), load1 = coredata(df$load1))
df2 <- data.frame(load2 = coredata(df2$load2))

df3 <- cbind(df, df2)
df3$load <- df3$load1 + df3$load2
df3 <- df3[,c(1,4)]
write.csv(df3, row.names = FALSE, 
          file = "./Competition/Data/NN.deseason.212.6768.csv")

```

##Neural_Network, for non-seasonal
p= 2,P = 0
```{r NN, non-seasonal,k(2,2)}
## remove seansaol
stl_decomposition <- stl(ts_load_daily_fill, s.window = "periodic", robust = TRUE)
seasonal_component <- stl_decomposition$time.series[, "seasonal"]
deseasonalized_data <- ts_load_daily_fill - seasonal_component

set.seed(6787)
NN_fit_deseasonalized_212 <- nnetar(deseasonalized_data, p = 2, P = 0)
NN_forecast_deseasonalized_212 <- forecast(NN_fit_deseasonalized_212, h = 59)

forecast_seasonal_component <- window(seasonal_component, start = c(2010, 1), end = c(2010, 59))


df <- data.frame(date = data_seq_YR, load1 = NN_forecast_deseasonalized_212$mean)
df2 <- data.frame(load2 = forecast_seasonal_component)
df3 <- cbind(df,df2) 

df <- data.frame(date = as.Date(data_seq_YR), load1 = coredata(df$load1))
df2 <- data.frame(load2 = coredata(df2$load2))

df3 <- cbind(df, df2)
df3$load <- df3$load1 + df3$load2
df3 <- df3[,c(1,4)]
write.csv(df3, row.names = FALSE, 
          file = "./Competition/Data/NN.deseason.p2.6768.csv")

```

##Arima+temp

```{r}
set.seed(6787)
deseasonalized_data_temp <- cbind(deseasonalized_data, ts_temp)

arima_fit_deseasonalized_temp <- auto.arima(deseasonalized_data_temp[, 1],
                                            xreg = as.matrix(deseasonalized_data_temp[, 3]),
                                            seasonal = FALSE)
xreg <- deseasonalized_data_temp[(nrow(deseasonalized_data_temp)-58):nrow(deseasonalized_data_temp), 3]

forecast_deseasonalized_T <- forecast(arima_fit_deseasonalized_temp, xreg = xreg, h = 59)


df <- data.frame(date = data_seq_YR, load1 = forecast_deseasonalized_T$mean)
df2 <- data.frame(load2 = forecast_seasonal_component)
df3 <- cbind(df,df2) 

df <- data.frame(date = as.Date(data_seq_YR), load1 = coredata(df$load1))
df2 <- data.frame(load2 = coredata(df2$load2))

df3 <- cbind(df, df2)
df3$load <- df3$load1 + df3$load2
df3 <- df3[,c(1,4)]
write.csv(df3, row.names = FALSE, 
          file = "./Competition/Data/NN.deseason.Temp.6768.csv")
```

##Neural_Network, k= 2,2,p=2
maybe need setseed
```{r Neural_Network(2,2)p=2}
set.seed(6787)
NN_fit_p2_YR <- nnetar(ts_load_daily,p=2,P=0,xreg=fourier(ts_load_daily, K=c(2,2)))

NN_forp2_YR <- forecast(NN_fit_p2_YR, h=59,xreg=fourier(ts_load_daily, 
                                          K=c(2,2),h=59))


#save
Neural.Network.p2.22.6787.YR_df <- data.frame(date = data_seq_YR, load = NN_forp2_YR$mean)

write.csv(Neural.Network.p2.22.6787.YR_df, row.names = FALSE, 
          file = "./Competition/Data/Neural.Network.Forecast.p2.22.seed6787.csv")


```
##Neural_Network, k= 2,2,p=1,P=1
maybe need setseed
```{r Neural_Network(2,2)p1P1}
set.seed(6787)
NN_fit_p1P1_YR <- nnetar(ts_load_daily,p=1,P=1,xreg=fourier(ts_load_daily, K=c(2,2)))

NN_for_p1P1_YR <- forecast(NN_fit_p1P1_YR, h=59,xreg=fourier(ts_load_daily, 
                                          K=c(2,2),h=59))


#save
Neural.Network.p1P1.22.6787.YR_df <- data.frame(date = data_seq_YR, load = NN_for_p1P1_YR$mean)

write.csv(Neural.Network.p1P1.22.6787.YR_df, row.names = FALSE, 
          file = "./Competition/Data/Neural.Network.Forecast.p1P1.22.seed6787.csv")

```
##Neural_Network, k= 2,2,p=1,P=4
maybe need setseed
```{r Neural_Network(2,2)p=1,P4}
set.seed(6787)
NN_fit_p1P4_YR <- nnetar(ts_load_daily,p=1,P=4,xreg=fourier(ts_load_daily, K=c(2,2)))

NN_for_p1P4_YR <- forecast(NN_fit_p1P4_YR, h=59,xreg=fourier(ts_load_daily, 
                                          K=c(2,2),h=59))


#save
Neural.Network.p1P4.22.6787.YR_df <- data.frame(date = data_seq_YR, load = NN_for_p1P4_YR$mean)

write.csv(Neural.Network.p1P4.22.6787.YR_df, row.names = FALSE, 
          file = "./Competition/Data/Neural.Network.Forecast.p1P4.22.seed6787.csv")

```
